### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_mediation.tex
new files for version 2.4.3
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1282 d8b0c59b-6544-0410-94d9-83a5658a058e
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1304 d8b0c59b-6544-0410-94d9-83a5658a058e

\newcommand\deliverableNumber{D2.4.2}
\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST platform}
\newcommand\deliverableTitleShort{Mediation amongst ontologies}
\newcommand\workpackageNumber{2}
\newcommand\workpackageTitle{Definition of Conceptual Model}
\newcommand\authorOne{Oszk\'{a}r Ambrus (NUIG)}
\newcommand\authorTwo{Albert Z\"{u}ndorf (UniKassel)}
\newcommand\authorThree{J\"{o}rn Friedrich Dreyer (UniKassel)}
\newcommand\authorFour{Sebasti\'{a}n Ortega (UPM)}
\newtheorem{example}{\emph{Example}}

@@

\pagenumbering{roman}

@@

%\begin{minipage}[p]{15cm}
    \begin{center}
        \includegraphics{images/FAST_logo}\\
        \vspace{1cm}
        {\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
        \vspace{4cm}
        {\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \textbf{\deliverableTitle}}\\
        \vspace{2cm}
        {\large \sffamily \authorOne} \\
        {\large \sffamily \authorTwo} \\
        {\large \sffamily \authorThree} \\
        {\large \sffamily \authorFour} \\
        \vspace{0.5cm}
        \vfill
        {\large \sffamily Date: 27/02/2010}\\
        \vspace{1cm}
        {\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
\rofoot{\small{\thepage}}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
                            \textcolor{white}{\textbf{Change description}}\\ \hline
% 0.1 & 09.05.2009 & \authorOne & Initial content \\ \hline
% 0.4 & 30.11.2009 & \authorOne & Sections 2 and 3 \\ \hline
% 0.5 & 07.01.2010 & \authorOne & New structure, Section 4 and 5 \\ \hline
% 0.7 & 03.02.2010 & \authorTwo & Section 3.5, 3.6, updated Section 5  \\ \hline
% 0.8 & 03.02.2010 & \authorOne & Integration and preliminary finalization  \\ \hline
% 0.9 & 03.02.2010 & NUIG, UniKassel & Corrections  \\ \hline
% 1.0 & 03.02.2010 & \authorOne & Integration and finalization  \\ \hline
%0.2 & 19.02.2009 & Oszkar Ambrus (NUIG) & content moved from wiki after internal review \\ \hline
%0.1 & 07.01.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
1.0 & 27.02.2009 & \authorOne & final version (D2.4.1) ready for external review \\ \hline
2.0 & 27.02.2010 & \authorOne & final version (D2.4.2) ready for external review \\ \hline
\end{tabular}
\end{small}

@@

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.2 is the second iteration of the FAST ontology matching approach. The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, this current interation presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\clearpage
\section*{Lists of Tables and Figures}
\listoftables
\listoffigures
\cleardoublepage
\pagenumbering{arabic}

@@

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:intro}
\subsection{Goal and Scope.}
FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones. Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets'', that rely on semantically enriched building blocks. In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties. These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary as the means to reconcile differences in the various conceptualisations.

We aim to integrate ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.  We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities. The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).

\subsubsection{Changes to Previous Version}

The first iteration, dealing with the theoretical foundations of ontology matching, is now contained in the Appendix, and the whole body of this document is newly added content, based on the comments of the last review. As such, it tackles design and development issues for integrating ontology matching within the FAST platform: it presents the overall strategy and deals with specific problems belonging to each phase of ontology matching in FAST.

\subsection{Structure of the Document.}
This document describes the approach of integrating ontology matching into the FAST environment. Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand. Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilites of the tool, and the testing procedure with the results. The remaing part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue. Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.

@@

\include{fast_d2_4_2_part1}

@@

For the last and final iteration of deliverable D2.4 we plan to evaluate real-life scenarios and finalize the implementation of ontology matching for the FAST platform.

@@

\include{fast_d2_4_2_part2}

\clearpage
\doublespacing
\appendix
% \section{Lists of Tables and Figures}
% %\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
% 
% \listoftables
% \listoffigures

>>>>

\newcommand\deliverableNumber{D2.4.3}
\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST platform}
\newcommand\deliverableTitleShort{Mediation amongst ontologies}
\newcommand\workpackageNumber{2}
\newcommand\workpackageTitle{Definition of Conceptual Model}
\newcommand\authorOne{Oszkar Ambrus (NUIG)}
\newcommand\authorTwo{Knud M\"{o}ller (NUIG)}
\newcommand\authorThree{Albert Z\"{u}ndorf (UniKassel)}
\newcommand\authorFour{J\"{o}rn Friedrich Dreyer (UniKassel)}
\newcommand\authorFive{Sebasti\'{a}n Ortega (UPM)}
\newtheorem{example}{\emph{Example}}

@@

%\pagenumbering{roman}
\pagenumbering{arabic}

@@

%\begin{minipage}[p]{15cm}
    \begin{center}
        \includegraphics{images/FAST_logo}\\
        \vspace{1cm}
        {\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
        \vspace{4cm}
        {\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \textbf{\deliverableTitle}}\\
        \vspace{2cm}
        {\large \sffamily \authorOne} \\
        {\large \sffamily \authorTwo} \\
        {\large \sffamily \authorThree} \\
        {\large \sffamily \authorFour} \\
        {\large \sffamily \authorFive} \\
        \vspace{0.5cm}
        \vfill
        {\large \sffamily Date: 27/02/2011}\\
        \vspace{1cm}
        {\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort\ $\bullet$ Page \thepage\ of \pageref{LastPage}}}
%\rofoot{\small{}}
%\rofoot{\small{\thepage}}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
                            \textcolor{white}{\textbf{Change description}}\\ \hline
% 0.1 & 09.05.2009 & \authorOne & Initial content \\ \hline
% 0.4 & 30.11.2009 & \authorOne & Sections 2 and 3 \\ \hline
% 0.5 & 07.01.2010 & \authorOne & New structure, Section 4 and 5 \\ \hline
% 0.7 & 03.02.2010 & \authorTwo & Section 3.5, 3.6, updated Section 5  \\ \hline
% 0.8 & 03.02.2010 & \authorOne & Integration and preliminary finalization  \\ \hline
% 0.9 & 03.02.2010 & NUIG, UniKassel & Corrections  \\ \hline
% 1.0 & 03.02.2010 & \authorOne & Integration and finalization  \\ \hline
%0.2 & 19.02.2009 & Oszkar Ambrus (NUIG) & content moved from wiki after internal review \\ \hline
%0.1 & 07.01.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
1.0 & 27.02.2009 & \authorOne & final version (D2.4.1) ready for external review \\ \hline
2.0 & 27.02.2010 & \authorOne & final version (D2.4.2) ready for external review \\ \hline
3.0 & 27.02.2011 & \authorTwo & final version (D2.4.3) ready for external review \\ \hline
\end{tabular}
\end{small}

@@

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.3 is the third iteration of the FAST ontology matching approach. The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for further analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, this current iteration presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour, \authorFive}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\clearpage
\section*{Lists of Tables and Figures}
\listoftables
\listoffigures
\cleardoublepage
%\pagenumbering{arabic}

@@

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:intro}
\subsection{Goal and Scope}
FAST is a visual programming platform allowing non-technical end users to build situational mashups, employing various underlying services and generating new ones. Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets'', that rely on semantically enriched building blocks. In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties. These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary as the means to reconcile differences in the various conceptualisations.

We aim to integrate a form of ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.  We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities. The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).

\subsubsection{Changes from Previous Version}

The following changes have been made with respect to the previous version of this deliverable, D2.4.2.~\cite{ambrus2010mediation}. A history of previous changes is kept in App.~\ref{sec:old_changes}.

\begin{itemize}
	\item In response to the previous review, we have extended and improved the text in order to better position ontology matching in FAST throughout the document.
	\item The previous D2.4.2 introduced the \emph{fact transformation tool}. At the time, this tool was an early stand-alone prototype. In the meantime, the fact tool has been properly integrated with the FAST tool chain. These improvements are reflected in both text and screenshots in Sects.~\ref{sec:fact_tool} and \ref{sec:creating_the_operator}. New content has been added to show how the output of the tool is used in screen development.
	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been made throughout the deliverable.
\end{itemize}

\subsection{Structure of the Document}
This document describes the approach of integrating ontology matching into the FAST environment. Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand. Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilities of the tool, and the testing procedure with the results. The remaining part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue. Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.

@@

\include{fast_d2_4_3_part1}

@@

For the last and final iteration of deliverable D2.4 we plan to evaluate real-life scenarios and finalise the implementation of ontology matching for the FAST platform.

@@

\include{fast_d2_4_3_part2}

\clearpage
\phantomsection
\section{Past Changes} % (fold)
\label{sec:old_changes}

For keeping a record of past changes, this section simply lists the changes that were made between the previous two versions of this deliverable, D2.4.1~\cite{ambrus2009mediation} and D2.4.2.~\cite{ambrus2010mediation}:

The first iteration, dealing with the theoretical foundations of ontology matching, had been moved to App.~\ref{sec:background}. The preceding main body of D2.4.2 document is newly added content, based on the comments of the last review (M24). As such, it tackles design and development issues for integrating ontology matching within the FAST platform: it presents the overall strategy and deals with specific problems belonging to each phase of ontology matching in FAST.

### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part1.tex
New screen dumps
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1284 d8b0c59b-6544-0410-94d9-83a5658a058e
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1304 d8b0c59b-6544-0410-94d9-83a5658a058e

@@

% \begin{abstract}
% The FAST environment allows users to graphically compose intelligent, i.e. semantically annotated gadgets from predefined building blocks and deploy them on various mashup platforms, thus enabling the interconnection of different systems and services. Since different parties involved in the creation of gadgets and building blocks conceptualise knowledge in different ontologies, using ontology matching as a means to reconcile differences in the various conceptualisations is a crucial issue.
% This paper discusses first steps in our effort to integrate ontology matching in an end-user-oriented environment such as FAST, using an e-commerce-related scenario. We evaluate a number of tools and approaches for solving different levels of complexity in ontology matching and 
% %, which we address in this paper. We present a scenario of interconnecting e-commerce web services within a FAST gadget, the ontologies created to represent those services and the alignment tool and language used to generate and represent the ontology mappings. The testing aims firstly, to find out whether and how state-of-the-art ontology matching makes it possible to mash-up different systems and secondly, to 
% define the direction of integrating ontology matching into the FAST environment.
% 
% \keywords{Ontology Matching, User-oriented, Mashups, E-commerce}
% \end{abstract}

@@

@@

FAST uses ontologies to conceptualise the underlying resources used by the different components.
Ontologies embody the fundamental vehicle for conceptualising data on semantic systems; they describe the context and semantic background of data that should be known to all agents using it~\cite{gruber93towards}.
However, different ontologies are often used to describe the same domain or cover the same scenario. This is also true for FAST, where gadget building blocks can originate from different providers, who might use different ontologies to describe them. To ensure interoperability, the task of ontology matching is therefore critical in FAST.

@@

The gadget life cycle in FAST has several phases and roles associated, as detailed in \cite{hoyer-fast}. Here, we list the ones relevant for the ontology matching tasks, in decreasing order of the measure in which knowledge about ontologies is required. Note that several roles can be played by the same actor.
\begin{enumerate}[(i)]
    \item The \textit{ontology engineer} creates the ontologies used to annotate services and data. This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST gadget building. The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions of a screen.
    \item Ontology matching is needed by the \textit{gadget developer} at the design-time of a screen-flow (gadget). Gadget developers use the GVS (Gadget Visual Storyboard) for building gadgets, where they can use the matching operators to combine screens having input and output conditions specifying resources annotated with different ontologies. No actual matching needs to be performed in this phase, but rather the possibility of matching needs to be determined (i.e., \emph{can} two screens A and B be combined?).
%the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
    %\item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
    \item The \textit{end-user} uses the final deployed gadget at run-time, but is unaware of the underlying resources and ontologies or the matching process. Only at run-time the actual mapping of instance data has to be performed.
\end{enumerate}

@@

@@

\subsection{Facts Transformation Tool}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitionsNoGlitch2.png}
    \caption{Definition of operator pre- / postconditions}
    \label{fig:FactsTransformationToolPortDefinitions}
\end{figure}

@@

\begin{figure}
  \begin{center}
\includegraphics[angle=90,width=0.8\linewidth]{images/FactsTransformationToolGVSWithTransformationRules.jpg}
    \caption{Interactive, rule based transforming of Person facts (photoshoped screen dump)}
    \label{fig:FactsTransformationToolTransformationRules}
  \end{center}
\end{figure}

\subsection{Creating the Operator}

Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool. For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions.

>>>>

%!TEX root = fast_d2_4_3_mediation.tex

@@

@@

In this section, we introduce some general terminology and definitions regarding ontology matching, as well as positioning ontology matching in FAST in terms of user roles.

@@

FAST uses ontologies to conceptualise resources and data used by the different components.
Ontologies embody the fundamental vehicle for conceptualising data on semantic systems; they describe the context and semantic background of data that should be known to all agents using it~\cite{gruber93towards}.
However, different ontologies are often used to describe the same domain or cover the same scenario. This is also true for FAST, where gadget building blocks can originate from different providers, who might use different ontologies to describe them. To ensure interoperability, the task of ontology matching is therefore important in FAST.

@@

As detailed in \cite{hoyer2009fast}, the gadget life cycle in FAST has several phases and roles associated. Here, we list the ones relevant for the ontology matching tasks, in decreasing order of the measure in which knowledge about ontologies is required. Note that several roles can be played by the same actor.
\begin{enumerate}[(i)]
    \item The \textit{ontology engineer} creates the ontologies used to annotate services and data. This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST gadget building. The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions of a screen. These three roles are the most critical in FAST with respect to ontology matching, and topics relevant to these roles will be discussed in detail in Sect.~\ref{sec:ontologyengineering}.
    \item Ontology matching is needed by the \textit{gadget developer} at the design-time of a screen-flow (gadget). Gadget developers use the GVS (Gadget Visual Storyboard) for building gadgets, where they can use the matching operators to combine screens having input and output conditions specifying resources annotated with different ontologies. No actual matching needs to be performed in this phase, but rather the possibility of matching needs to be determined (i.e., \emph{can} two screens A and B be combined?).
%the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
Section~\ref{sec:gadgetbuilding} outlines how the gadget developer if affected by ontology matching.
    %\item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
    \item The \textit{end-user} uses the final deployed gadget at run-time, but is unaware of the underlying resources and ontologies or the matching process. Only at run-time the actual mapping of instance data has to be performed.
This aspect of ontology matching is briefly discussed in Sect.~\ref{sec:runtime}.
\end{enumerate}

@@

In the FAST lifecycle, the ontology engineering phase is the most relevant for the topic of this deliverable. Relevant roles are the ontology engineer, who performs the main work in this phase, as well as the the resource developer and the screen developer, who are directly affected.

@@

\subsection{Fact Transformation Tool}
\label{sec:fact_tool}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{images/GVSMediationToolPreconditions.png}
    \caption{Definition of operator pre- / postconditions}
    \label{fig:FactsTransformationToolPortDefinitions}
\end{figure}

@@

\begin{figure}
  \begin{center}
\includegraphics[width=\linewidth]{images/FactsTransformationToolGVSWithTransformationRules.png}
    \caption{Interactive, rule-based transformation of Person facts}
    \label{fig:FactsTransformationToolTransformationRules}
  \end{center}
\end{figure}

\subsection{Creating the Operator}
\label{sec:creating_the_operator}

Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST GVS tool. For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions. This is shown in Fig.~\ref{fig:operator_in_screen_builder}, where the newly created operator for mediating between Person and Customer concepts (i.e., transforming instances of one concepts into instances of the other) has been saved and is now available in the GVS screen builder interface.

\begin{figure}[h]
  \begin{center}
\includegraphics[width=.6\linewidth]{images/operator_closeup}
    \caption{A newly created operator in the GVS screen builder}
    \label{fig:operator_in_screen_builder}
  \end{center}
\end{figure}
